Stack Safety as a Security Property

### 2020-07-07 main.tex
More on capabilities and StkTokens

@@

StkTokens \citep{Skorstengaard+19} protects the stack by enforcing local state
encapsulation and well-bracketed control flow. It does so by defining a calling
convention that makes use of linear capabilities for stack and return pointers
operating on a single shared stack, and by requiring that protected components
avoid compromising their own security by following certain simple rules.
Informally, local state encapsulation is defined by restricting accesses to the
range of memory allocated to the current stack frame, and well-bracketed control
flow by only allowing returns from the topmost frame to the immediately adjacent
frame below. Formally, these properties are built into a capability machine
defined by an overlay semantics, which is proven fully abstract with respect to
a capability machine that replaces those abstract operations with their calling
convention.

\rb{There are quite a few parallels at various levels, although the definitions
  are overall more implicit and not given formally. A limiting factor is the
  lack of linear capabilities in existing models like CHERI---although these
  could be implemented by a micropolicy.}

>>>>

\rb{Maybe provide a full definition of (eager) stack safety, here or at the end
  of the section.}

@@

CHERI \rb{cite}

\citet{Skorstengaard+19b} introduce a calling convention to preserve local state
encapsulation and well-bracketed control flow by the use of local capabilities.
Although the required hardware support is readily available in capability
machines like CHERI, this technique incurs potentially prohibitive performance
costs, because it requires the unused part of the stack to be cleared whenever a
security boundary is crossed. In the discussion, ``while [the authors] claim
that [their] calling convention enforces control-flow correctness, [they] do not
prove a general theorem that shows this, because it is not clear what such a
theorem should look like,'' noting that the correctness property enjoyed by
their technique ``is not made very explicit.''

StkTokens \citep{Skorstengaard+19} protects the stack by enforcing local state
encapsulation and well-bracketed control flow. It does so by defining a calling
convention that makes use of linear capabilities for stack and return pointers
operating on a single shared stack, and by requiring that protected components
avoid compromising their own security by following certain simple rules.
Informally, local state encapsulation is defined by restricting accesses to the
range of memory allocated to the current stack frame, and well-bracketed control
flow by only allowing returns from the topmost frame to the immediately adjacent
frame below. Formally, it improves on \citet{Skorstengaard+19b} by building
those properties into the semantics of a capability machine with a built-in call
stack and call and return instructions, which is proven fully abstract with
respect to a more concrete capability machine that replaces those
pseudo-instructions with their calling convention.

We are confident \rb{how strong a claim to make, with substantiation} that
StkTokens satisfies the eager stack safety property, and even the stronger eager
testing property \rb{refs}. The main limitation of StkTokens is its reliance on
linear capabilities, as it is unclear now they could be added to practical
capability machines, especially in terms of efficiency.
%
\rb{On the other hand, we know that these can be implemented as micropolicies,
  cf. linear return capabilities in yannis-report (what about performance?)}

\rb{There are quite a few parallels at various levels, although the definitions
  are overall more implicit and not given formally.}
